Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-06-16 performance_comparison.tex
[mg] fixes

%\subsection{Open-loop vs closed-loop robust planning}
\begin{figure}[b!]
	\centering
	\includegraphics{Fig/olcl_traj_bands.pdf}
	\caption{Back-off envelopes for the open-loop (red) and closed-loop (cyan) robust planners. Each shaded band represents the area swept by the back-off-extended vehicle and must lie entirely within the actual track boundaries.
		The open-loop approach yields a wider envelope due to worst-case growth assumptions, whereas the closed-loop design, aided by time-varying LQR controller, maintains a narrower margin.}
	\label{fig:olcl_traj_bands}
\end{figure}
Finally, we compare the influence of the back-off terms on the open-loop and closed-loop trajectories, in the robustified TLC scenario with the high $\ga^\textrm{TLC}=3.0$ value for better visualization. As before, the open-loop planning is carried out with a prediction horizon of $H=4$ steps.
The back-off term $\be^{TLC}$ in Eq.~\eqref{eq:TLC} admits an interesting geometrical interpretation: it corresponds to the amount by which the track boundaries are (not uniformly) tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width.

%According to the latter interpretation, Fig.~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle. Ultimately, the TLC enforces precisely that these bands remain entirely within the actual track limits. The comparison between the two regions illustrates how the additional conservativeness of the open-loop approach stems from larger back-off terms, i.e., a wider band that must stay within the track boundaries. The closed-loop method, on the other hand, benefits from the LQR-based feedback, which allows it to maintain smaller back-off terms and thus a narrower band, requiring a reduced safety margin from the track limits.

@@

\subsection{Simulations with noise realization}
To provide an empirical validation of the closed-loop robustified planning we simulated the evolution of the controlled vehicle with 100 random initial conditions and included noise realizations. The random values for the initial state and the noise follow a Gaussian distribution with covariance matrices $\bP_0$ and $\bQ$, in both cases, as introduced in Eq.~\eqref{eq:dP}. This analysis is not applicable the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
%For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.
\begin{figure}[t!]
	\centering
	\includegraphics{Fig/olcl_traj_strings.pdf}
	\caption{100 simulations (each panel) with random initial conditions and Gaussian noise realizations. In the left panel, the reference trajectory (black curve) is the non-robust one, tracked using an LQR controller designed around it as per Step 2 of Subsection~\ref{sec:LQR}. In the right panel, the reference (black curve) is a robust trajectory obtained using the closed-loop optimization method as per Step 3 of Subsection~\ref{sec:CLrobustified}, which is tracked using the corresponding optimized controller.}
	\label{fig:traj_strings}
\end{figure}

>>>>

%\subsection{Open-loop vs closed-loop robust planning}
\begin{figure}[b!]
	\centering
	\includegraphics{Fig/olcl_traj_bands.pdf}
	\caption{Back-off envelopes for the open-loop (red) and closed-loop (cyan) robust planners. Each shaded band represents the area swept by the back-off-extended vehicle and must lie entirely within the actual track boundaries.
		The open-loop approach yields a wider envelope due to worst-case growth assumptions, whereas the closed-loop design, aided by time-varying LQR controller, maintains a narrower margin.}
	\label{fig:olcl_traj_bands}
\end{figure}

Finally, we compare the influence of the back-off terms on the open-loop and closed-loop trajectories, in the robustified TLC scenario with the high $\ga^\textrm{TLC}=3.0$ value for better visualization. As before, the open-loop planning is carried out with a prediction horizon of $H=4$ steps.
The back-off term $\be^{TLC}$ in Eq.~\eqref{eq:TLC} admits an interesting geometrical interpretation: it corresponds to the amount by which the track boundaries are (not uniformly) tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width.

According to the latter interpretation, Fig.~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle. Ultimately, the TLC enforces precisely that these bands remain entirely within the actual track limits. The comparison between the two regions illustrates how the additional conservativeness of the open-loop approach stems from larger back-off terms, i.e., a wider band that must stay within the track boundaries. The closed-loop method, on the other hand, benefits from the LQR-based feedback, which allows it to maintain smaller back-off terms and thus a narrower band, requiring a reduced safety margin from the track limits.

@@

\subsection{Simulations with noise realization}
\begin{figure}[t!]
	\centering
	\includegraphics{Fig/olcl_traj_strings.pdf}
	\caption{100 simulations (each panel) with random initial conditions and Gaussian noise realizations. In the left panel, the reference trajectory (black curve) is the non-robust one, tracked using an LQR controller designed around it as per Step 2 of Subsection~\ref{sec:LQR}. In the right panel, the reference (black curve) is a robust trajectory obtained using the closed-loop optimization method as per Step 3 of Subsection~\ref{sec:CLrobustified}, which is tracked using the corresponding optimized controller.}
	\label{fig:traj_strings}
\end{figure}
To provide an empirical validation of the closed-loop robustified planning we simulated the evolution of the controlled vehicle with 100 random initial conditions and included noise realizations. The random values for the initial state and the noise follow a Gaussian distribution with covariance matrices $\bP_0$ and $\bQ$, in both cases, as introduced in Eq.~\eqref{eq:dP}. This analysis is not applicable the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
%For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.
